### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@3ef21deae4d7de31ea2b5a3daee2d1c1559c4e45 🚀

However, we find that the global pattern of betacoronavirus phylogenetic
distinctiveness is quite distinct from both bat host richness and
phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
contrast to the sparsity of Neotropical betacoronavirus hosts, South and
Central America have the most evolutionary distinct hosts \emph{and}
viruses, followed by secondary hotspots in southeast Asia and the Rift
Valley region have mostly distinct viruses. Some degree of sampling bias
may contribute to these patterns: for example, the Neotropics are one of
the places where the fewest bat betacoronavirus sequences have been
generated,\textsuperscript{\textbf{Worobey2022HuaMar?},\textbf{Temmam2022BatCor?},\textbf{Boni2020EvoOri?}}
resulting in a sparser phylogenetic tree, and artificially inflating
distinctiveness; conversely, disproportionate research effort in eastern
China\textsuperscript{\textbf{Cohen2022SamStr?}} may have led to a more
complete inventory of the local diversity of coronaviruses, again
inflating these metrics relative to underlying patterns. Even accounting
for these potential biases, though, there is obvious heterogeneity in
betacoronavirus evolutionary distinctiveness that is distinct from
overall bat diversity.

Overall, these patterns recapitulate the evolutionary history of both
the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
bats (Rhinolophidae) include the reservoirs of the SARS-like viruses
(subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
been of the greatest interest to
researchers\textsuperscript{\textbf{Latinne2020OriCro?}} (and so have
been sampled most
intensively).\textsuperscript{\textbf{Cohen2022SamStr?}} The hotspots of
host richness and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---are almost entirely driven by viral adaptive radiation
through host switching within this
clade\textsuperscript{\textbf{Becker2022OptPre?},\textbf{Ruiz-Aravena2022EcoEvo?}}.
In contrast, the Neotropical hotspot of viral distinctiveness is driven
by isolation by host vicariance. Out of the four main groups of
betacoronaviruses, only merbecoviruses have been found in animals in the
Americas--- an introduction that is generally presumed to be
ancient.\textsuperscript{\textbf{Ruiz-Aravena2022EcoEvo?},\textbf{Olival2020PosRev?}}
While comparatively understudied, New World merbecoviruses have been
found in the ghost-faced bats (Mormoopidae), Neotropical leaf-nosed bats
(Phyllostomidae), and free-tailed bats
(Molossidae).\textsuperscript{\textbf{Anthony2013CorBat?},\textbf{Goes2013NovBat?},\textbf{Goes2016GenDiv?},\textbf{Brandao2008CorDet?}}
The former two groups and a clade of the latter are endemic to the
Neotropics, while the explosive adaptive radiations of the phyllostomids
are responsible for the hotspot of bat diversity in the
Amazon.\textsuperscript{\textbf{Ammerman2012FirMol?}} Together, these
clades of New World bats play host to a distinct regime of
betacoronavirus coevolution.

@@

Bats---and the spillover of their viruses---are also sensitive to
anthropogenic factors others than climate change, including
deforestation and other kinds of habitat loss, increased stress, and
greater contact with potential bridge hosts like domesticated
species.\textsuperscript{\textbf{Alves2018GeoVar?},\textbf{Treitler2016EffLoc?},\textbf{Rulli2021LanCha?},\textbf{Mendenhall2014PreBio?}}
This represents a challenge for both conservation strategies and
pandemic prevention,\textsuperscript{\textbf{Amman2011InvRol?}} but
identifying areas at risk, and protecting the health of bats and
ecosystems within those zones, can be a win-win intervention for
both.\textsuperscript{\textbf{Hopkins2021HowIde?},\textbf{Plowright2021LanUse?},\textbf{OHHLEP2022OneHea?}}
As we scale these predictions down in space to finer spatial resolutions
to guide public health actions,\textsuperscript{21} the incorporation of
human activity predictors will become more
importyant.\textsuperscript{22}

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{\textbf{Albery2020PreGlo?}} This network stores
pairwise values of viral community similarity, measured for all hosts
(to maintain consistency with teh phylogenetic diversity measure) across
all viruses; therefore, we consider that it accounts for some overall
similarity in the way hosts deal with viruses, and not only
betacoronaviruses. There is empirical evidence that capacity for
cross-species transmission even between divergent species is generally
high,\textsuperscript{23} especially for
beta-coronaviruses.\textsuperscript{13} To project viral sharing values
into a single value for every pixel, we averaged the pairwise scores.
High values of the average sharing propensity means that this specific
extant bat assemblage is likely to be proficient at exchanging viruses.

@@

\leavevmode\hypertarget{ref-Muylaert2022Present}{}%
\CSLLeftMargin{21. }
\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
distribution of bat hosts of sarbecoviruses: Implications for
conservation and public health. \emph{Proceedings of the Royal Society
B: Biological Sciences} \textbf{289}, 20220397 (2022).}

\leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
\CSLLeftMargin{22. }
\CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
re-emerging viral infectious diseases. \emph{Microbes and Infection}
\textbf{8}, 905--916 (2006).}

\leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
\CSLLeftMargin{23. }
\CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
homogenous among taxonomic orders of mammalian and avian reservoir
hosts. \emph{Proceedings of the National Academy of Sciences}
\textbf{117}, 9423--9430 (2020).}

>>>>

However, we find that the global pattern of betacoronavirus phylogenetic
distinctiveness is quite distinct from both bat host richness and
phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
contrast to the sparsity of Neotropical betacoronavirus hosts, South and
Central America have the most evolutionary distinct hosts \emph{and}
viruses, followed by secondary hotspots in southeast Asia and the Rift
Valley region have mostly distinct viruses. Some degree of sampling bias
may contribute to these patterns: for example, the Neotropics are one of
the places where the fewest bat betacoronavirus sequences have been
generated,\textsuperscript{21--23} resulting in a sparser phylogenetic
tree, and artificially inflating distinctiveness; conversely,
disproportionate research effort in eastern China\textsuperscript{24}
may have led to a more complete inventory of the local diversity of
coronaviruses, again inflating these metrics relative to underlying
patterns. Even accounting for these potential biases, though, there is
obvious heterogeneity in betacoronavirus evolutionary distinctiveness
that is distinct from overall bat diversity.

Overall, these patterns recapitulate the evolutionary history of both
the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
bats (Rhinolophidae) include the reservoirs of the SARS-like viruses
(subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
been of the greatest interest to
researchers\textsuperscript{\textbf{Latinne2020Origins?}} (and so have
been sampled most intensively;).\textsuperscript{24} The hotspots of
host richness and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---are almost entirely driven by viral adaptive radiation
through host switching within this clade\textsuperscript{12,20}. In
contrast, the Neotropical hotspot of viral distinctiveness is driven by
isolation by host vicariance. Out of the four main groups of
betacoronaviruses, only merbecoviruses have been found in animals in the
Americas--- an introduction that is generally presumed to be
ancient.\textsuperscript{12,25} While comparatively understudied, New
World merbecoviruses have been found in the ghost-faced bats
(Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
free-tailed bats (Molossidae).\textsuperscript{14--17} The former two
groups and a clade of the latter are endemic to the Neotropics, while
the explosive adaptive radiations of the phyllostomids are responsible
for the hotspot of bat diversity in the Amazon.\textsuperscript{26}
Together, these clades of New World bats play host to a distinct regime
of betacoronavirus coevolution.

@@

Bats---and the spillover of their viruses---are also sensitive to
anthropogenic factors others than climate change, including
deforestation and other kinds of habitat loss, increased stress, and
greater contact with potential bridge hosts like domesticated
species.\textsuperscript{\textbf{Alves2018GeoVar?},\textbf{Treitler2016EffLoc?},\textbf{Rulli2021LanCha?},\textbf{Mendenhall2014PreBio?}}
This represents a challenge for both conservation strategies and
pandemic prevention,\textsuperscript{\textbf{Amman2011InvRol?}} but
identifying areas at risk, and protecting the health of bats and
ecosystems within those zones, can be a win-win intervention for
both.\textsuperscript{\textbf{Hopkins2021HowIde?},\textbf{Plowright2021LanUse?},\textbf{OHHLEP2022OneHea?}}
As we scale these predictions down in space to finer spatial resolutions
to guide public health actions,\textsuperscript{27} the incorporation of
human activity predictors will become more
importyant.\textsuperscript{28}

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{\textbf{Albery2020PreGlo?}} This network stores
pairwise values of viral community similarity, measured for all hosts
(to maintain consistency with teh phylogenetic diversity measure) across
all viruses; therefore, we consider that it accounts for some overall
similarity in the way hosts deal with viruses, and not only
betacoronaviruses. There is empirical evidence that capacity for
cross-species transmission even between divergent species is generally
high,\textsuperscript{29} especially for
beta-coronaviruses.\textsuperscript{13} To project viral sharing values
into a single value for every pixel, we averaged the pairwise scores.
High values of the average sharing propensity means that this specific
extant bat assemblage is likely to be proficient at exchanging viruses.

@@

\leavevmode\hypertarget{ref-Worobey2022Huanan}{}%
\CSLLeftMargin{21. }
\CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022)
doi:\href{https://doi.org/10.5281/zenodo.6299600}{10.5281/zenodo.6299600}.}

\leavevmode\hypertarget{ref-Temmam2022Bat}{}%
\CSLLeftMargin{22. }
\CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
SARS-CoV-2 and infectious for human cells. \emph{Nature} \textbf{604},
330--336 (2022).}

\leavevmode\hypertarget{ref-Boni2020Evolutionary}{}%
\CSLLeftMargin{23. }
\CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
SARS-CoV-2 sarbecovirus lineage responsible for the COVID-19 pandemic.
\emph{Nature microbiology} \textbf{5}, 1408--1417 (2020).}

\leavevmode\hypertarget{ref-Cohen2022Sampling}{}%
\CSLLeftMargin{24. }
\CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
Becker, D. J. Sampling strategies and pre-pandemic surveillance gaps for
bat coronaviruses. 2022.06.15.496296 (2022)
doi:\href{https://doi.org/10.1101/2022.06.15.496296}{10.1101/2022.06.15.496296}.}

\leavevmode\hypertarget{ref-Olival2020Possibility}{}%
\CSLLeftMargin{25. }
\CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
zoonotic transmission of SARS-CoV-2 to free-ranging wildlife: A case
study of bats. \emph{PLoS pathogens} \textbf{16}, e1008758 (2020).}

\leavevmode\hypertarget{ref-Ammerman2012First}{}%
\CSLLeftMargin{26. }
\CSLRightInline{Ammerman, L. K., Lee, D. N. \& Tipps, T. M. First
molecular phylogenetic insights into the evolution of free-tailed bats
in the subfamily Molossinae (Molossidae, Chiroptera). \emph{Journal of
Mammalogy} \textbf{93}, 12--28 (2012).}

\leavevmode\hypertarget{ref-Muylaert2022Present}{}%
\CSLLeftMargin{27. }
\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
distribution of bat hosts of sarbecoviruses: Implications for
conservation and public health. \emph{Proceedings of the Royal Society
B: Biological Sciences} \textbf{289}, 20220397 (2022).}

\leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
\CSLLeftMargin{28. }
\CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
re-emerging viral infectious diseases. \emph{Microbes and Infection}
\textbf{8}, 905--916 (2006).}

\leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
\CSLLeftMargin{29. }
\CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
homogenous among taxonomic orders of mammalian and avian reservoir
hosts. \emph{Proceedings of the National Academy of Sciences}
\textbf{117}, 9423--9430 (2020).}
